StreamReduce: A Data Stream Processor with MapReduce Parallelism

### 2012-05-11 paper/paper.tex
Add UAP completion comment to acknowledgments
Better title
Another editing round

\title{StreamReduce: Processing Data Streams\\and Avoiding Work}
\author{Ryan Lopopolo\\
Massachusetts Institute of Technology\\
\href{mailto:lopopolo@mit.edu}{\texttt{lopopolo@mit.edu}}}

@@

Another limitation of MapReduce is batch processing. How does one
incorporate new data into an already analyzed task? MapReduce simply says to rerun
the job with the new data included. Others have solved this problem by batching data
into chunks and continuously spawning MapReduce jobs to incorporate the new data
and incrementally storing updates in an external database instead of relying on
the file system for enforcing the synchronization barrier.

@@

Dynamic Knobs~\cite{Hoffmann:2011:DKR:1950365.1950390} describes a mechanism to
explore the
accuracy-performance tradeoff space, dynamically adjusting program tuning parameters in
response
to changes in the execution environment. This allows programs to do things such as
utilize more resources if they become available, use less power if power rates spike,
or use less accurate computations in the face of an impending time constraint.

@@

\section{Design}
\label{sec:design}
StreamReduce is a distributed system for processing streams that can choose to kill tasks or
limit the amount of work it performs per task. Its design is modular and there are five
different types of nodes in a StreamReduce cluster:
\begin{itemize}
  \item
    \textbf{Spout}: A node external to the cluster that is a source of streaming data.
  \item
    \textbf{Fetcher}: Communicates with spouts to provide data for workers.
  \item
    \textbf{Worker}: Processes data returned by fetchers. They are the source of parallel
computation
    in the cluster.
  \item
    \textbf{Collector}: Aggregates results from all workers and makes the result accessible via
an API.
  \item
    \textbf{Master}: Manages the cluster by bringing nodes online to complete jobs. It is
responsible
    for all internode communication and storing to-be-processed data in several queues.
\end{itemize}
Figure~\ref{fig:systemdiagram} shows a system diagram of StreamReduce.

@@

\subsection{Spout}
Spouts are network-connected machines that generate data for a StreamReduce cluster to
process.
Examples include a machine that tails a log file or an API like the Twitter tweet
stream. Fetchers depend on spouts as a source of streaming data.
\subsection{Fetcher}
Fetchers are responsible for communicating with spouts and returning data for the
workers to process. This level of indirection allows for a variety of different
configurations: A spout can emit data at such a high rate that using multiple fetchers
returns more data to the cluster, multiple fetchers can each talk to a different spout,
for example tailing logs from multiple machines, or fetchers can be set up redundantly
to ensure that every datum emitted by a spout is processed.

@@

Workers are responsible for fulfilling the main design goals of StreamReduce. Workers may
choose
to kill a task preemptively or, if it does not complete before a timeout, greedily. Additionally,
workers may choose among a variety of routines for performing computation on a datum.
Routines are ranked on a [0,1] scale based on the estimated fraction of total work it
requires. This rank is a proxy for the accuracy of a routine.

As it runs, a worker maintains an internal metric that estimates the fraction of work
it has performed over all data it has seen. Workers use this metric when choosing which
computation routine to run for a datum. A worker will choose the lowest ranked routine
such that its estimated fraction of work is above a configured threshold.
The total work estimator is used as a proxy for accuracy when a worker updates
this metric. This metric also takes into account the number of tasks it has killed
for any reason.

Because one of StreamReduce's goals is to process (potentially infinite) streams, workers do
not
save their progress to disk and are forgetful; once the master queries a worker for its
progress and the worker successfully returns it, the worker clears all of its state and
continues processing. As a consequence, workers are easy to add or remove from the worker
pool because they don't need to be initialized with long-term job state. Additionally, as
long as the master polls workers frequently enough, only a small
fraction of the total work is lost upon worker failure.
The master queries workers every 100ms by default.
\subsection{Collector}
Collector nodes aggregate the results of all workers. Each collector acts as a reducer would
in the MapReduce framework if MapReduce used a single reducer for all of the mappers.
Collectors differ from MapReduce reducers in that they receive their data piecewise, as it
is produced from workers, as opposed to after all mappers have completed. This limits the
types of stateless aggregation collectors can do to online operations. (Of course, a collector
can choose to retain copies of individual worker results and re-reduce them each time more
results are received, but this is not in the spirit of the framework.)

@@

The master is the only well-known node in the cluster. Other nodes in the network
exclusively
communicate with the master. The master and other nodes communicate with each other by
issuing
HTTP requests to a simple server on each node. I decided to make use of a
library~\cite{sinatrarb}
providing the server functionality because there was a low barrier to getting inter-node
communication functioning. This design choice presents some problems and will be
discussed
further in Section~\ref{sec:discussion}.

Another consequence of the master being the only well-known node is that all data in the
cluster must flow through it. The master maintains two queues per job for buffering job
progress from fetchers and workers and periodically pushes this state out to workers and
collectors, respectively.

@@

\begin{figure}
\includegraphics[width=\linewidth]{sr-topology.png}
\caption{Topology of the test StreamReduce cluster used in the word count case study. Each
square box represents a single machine. Each rounded box represents a StreamReduce role.}
\label{fig:clusterDiagram}
\end{figure}

@@

To run the job, I had to create a spout and find a source of data. For data, I acquired a
Wikipedia dump of the Featured Articles category. I used Wikipedia's xml
exporter~\cite{wikipedia:special:export} on January~9,~2012.

To serve the data, I transformed it into a YAML file containing a hash of
\begin{center}
$\texttt{article\_id}\rightarrow\texttt{fulltext}$.
\end{center}
To store fulltext, I base64 encoded it and
deflated it using zlib. This requires workers to inflate and decode the fulltext before they
can process it further. The server reads from this YAML file and returns the resulting hash as
JSON.

@@

@@

>>>>

\title{StreamReduce: A Data Stream Processor with MapReduce Parallelism}
\author{Ryan Lopopolo\\
Massachusetts Institute of Technology\\
\href{mailto:lopopolo@mit.edu}{\texttt{lopopolo@mit.edu}}}

@@

Another limitation of MapReduce is batch processing. How does one
incorporate new data into an already analyzed task? MapReduce simply says to rerun
the job with the new data included. Others have solved this problem by batching data
into chunks, continuously spawning MapReduce jobs to incorporate the new data,
and incrementally storing updates in an external database instead of relying on
the file system for enforcing the synchronization barrier.

@@

Dynamic Knobs~\cite{Hoffmann:2011:DKR:1950365.1950390} describes a mechanism to
explore the
accuracy--performance tradeoff space, dynamically adjusting program tuning parameters in
response
to changes in the execution environment. This allows programs to do things such as
utilize more resources if they become available, use less power if power rates spike,
or use less accurate computations in the face of an impending time constraint.

@@

\section{Design}
\label{sec:design}
StreamReduce is a distributed system for processing streams that can choose to kill tasks or
limit the amount of work it performs per task. Its design is modular and there are five
different types of nodes in a StreamReduce cluster:
\begin{description}
  \item[Spout] A node external to the cluster that is a source of streaming data.
  \item[Fetcher] Communicates with spouts to provide data for workers.
  \item[Worker] Processes data returned by fetchers. They are the source of parallel
computation
    in the cluster.
  \item[Collector] Aggregates results from all workers and makes the result accessible via
an API.
  \item[Master] Manages the cluster by bringing nodes online to complete jobs. It is
responsible
    for all internode communication and storing to-be-processed data in several queues.
\end{description}
Figure~\ref{fig:systemdiagram} shows a system diagram of a StreamReduce cluster.

@@

\subsection{Spout}
Spouts are network-connected machines that generate data for a StreamReduce cluster to
process.
Examples include a machine that tails a log file or an API like the Twitter tweet
stream~\cite{twitter}. Fetchers depend on spouts as a source of streaming data.
\subsection{Fetcher}
Fetchers are responsible for communicating with spouts and returning data for the
workers to process. This level of indirection allows for a variety of different
configurations: A spout can emit data at such a high rate that using multiple fetchers
returns more data to the cluster; multiple fetchers can each talk to a different spout,
for example tailing logs from multiple machines; or fetchers can be set up redundantly
to ensure that every datum emitted by a spout is processed.

@@

Workers are responsible for fulfilling the main design goals of StreamReduce. Workers may
choose
to kill a task preemptively or, if it does not complete before a timeout, greedily. Additionally,
workers may choose among a variety of routines for performing computation on a datum.
Routines are ranked on a $[0,1]$ scale based on the estimated fraction of total work it
requires. This rank is a proxy for the accuracy of a routine.

As it runs, a worker maintains an internal metric that estimates the fraction of work
it has performed over all data it has seen. Workers use this metric when choosing which
computation routine to run for a datum. A worker will choose the lowest ranked routine
such that its estimated fraction of work is above a configured threshold.
The total work estimator is used as a proxy for accuracy when a worker updates
this metric. This metric also takes into account the number of tasks a worker has killed
for any reason.

Because one of StreamReduce's goals is to process potentially infinite streams, workers do
not
save their progress to disk and are forgetful; once the master queries a worker for its
progress and the worker successfully returns it, the worker clears all of its state and
continues processing. As a consequence, workers are easy to add or remove from the worker
pool because they don't need to be initialized with long-term job state. Additionally, as
long as the master polls workers frequently enough, only a small
fraction of the total work is lost upon worker failure.
The master queries workers every 100ms by default.
\subsection{Collector}
Collector nodes aggregate the results of all workers. Each collector acts as a reducer would
in the MapReduce framework if MapReduce used a single reducer for all of the mappers.
Collectors differ from MapReduce reducers in that they receive their data piecewise, as it
is produced from workers, as opposed to after all mappers have completed. This limits the
types of stateless aggregation collectors can do to online operations. (Of course, a collector
can choose to retain copies of individual worker results and re-reduce them each time more
results are received, but this is not in the spirit of the framework.)

@@

The master is the only well-known node in the cluster. Other nodes in the network
exclusively
communicate with the master. The master and other nodes communicate with each other by
issuing
HTTP requests to a simple server on each node. I decided to make use of a
library~\cite{sinatrarb}
providing the server functionality because there was a low barrier to getting internode
communication functioning. This design choice presents some problems and will be
discussed
further in Section~\ref{sec:discussion}.

Another consequence of the master being the only well-known node is that all data in the
cluster must flow through it. The master maintains two queues per job for buffering job
progress from fetchers, the data queue, and workers, the result queue, and periodically
pushes this state out to workers and collectors, respectively.

@@

\begin{figure}[t]
\includegraphics[width=\linewidth]{sr-topology.png}
\caption{Topology of the test StreamReduce cluster used in the word count case study. Each
square box represents a single machine. Each rounded box represents a StreamReduce role.}
\label{fig:clusterDiagram}
\end{figure}

@@

To run the job, I had to create a spout and find a source of data. For data, I acquired a
Wikipedia dump of the Featured Articles category. I used Wikipedia's XML
exporter~\cite{wikipedia:special:export} on January~9,~2012.

To serve the data, I transformed it into a YAML file containing a hash of
\begin{center}
$\texttt{article\_id}\rightarrow\texttt{fulltext}$.
\end{center}

To store fulltext, I base64 encoded it and
deflated it using zlib. This requires workers to inflate and decode the fulltext before they
can process it further. The server reads from this YAML file and returns the resulting hash as
JSON.

@@

Running the perfect computation on one node took 1088 seconds. Compared to the fastest time
in Table~\ref{table:runtime} using 4 workers, this is over a~700\% speedup. There is much to
be gained by processing data in parallel.

@@

This paper was written in fulfillment of the MIT Course VI Undergraduate Advanced Project
requirement.
